---
title: "Tracking White-Tailed Deer Movement Patterns in the Duke Forest using Camera Traps"
author: "Sophie Moyer, Grace Randall, Ale Rios, Katie Tack"
date: "`r Sys.Date()`"
output: pdf_document
toc: true
---

## Content: 

The data used in this project is of animal sightings on the trail cameras around the Duke forest and the spatial data of the Duke forest;it was collected as part of an ongoing project monitoring the white-tail deer population.

## Hypotheses: 

1. We aim to determine the population size of white-tail deer in the Duke forest.
1. We aim to measure which areas of the forest were most frequented relative to development. (Source: Oleniacz, L. (2022, October 27). Scientists track triangle deer to learn how they deal with development. NC State University. Retrieved from https://news.ncsu.edu/2022/10/scientists-track-triangle-deer-to-learn-how-they-deal-with-development/)

## Background: 

To conduct a deer survey with trail cameras, one camera must be placed every 160 or so acres. (Source: Thomas Jr., L. (2012, April 19). How to run a trail-camera survey. Quality Deer Management Association.)
The Duke forest is 7,000 acres.(Source: Duke University. (n.d.). Duke Forest – Teaching and Research Laboratory. Retrieved from https://dukeforest.duke.edu)
There are 30 trail cameras in the Duke forest along known migration paths.

## Setup & Data Wrangling
Section created by Alejandra Rios

### Importing & Working Directory

```{r Import Packages, warning=FALSE, message=FALSE}
#import packages
library(tidyverse)
library(lubridate)
library(viridis)
library(here)
library(rvest);
library(dataRetrieval)
library(dplyr)
library(readr)
library(stringr)
library(sf); 
library(mapview); mapviewOptions(fgb = FALSE)
library(RColorBrewer)
```

```{r Get Working Directory, warning=FALSE, message=FALSE}
#get working directory
here()
```

```{r Read in Files, warning=FALSE, message=FALSE}
#read excel file 
trailcam_csv <- read.csv(here("sequences.csv"))
```

---

### Date & Time Objects

```{r Date as Objects, warning=FALSE, message=FALSE}
#date as objects
trailcam_csv$start_time <- ymd_hms(trailcam_csv$start_time)
trailcam_csv$end_time <- ymd_hms(trailcam_csv$end_time)
```

```{r Separate Date & Time, warning=FALSE, message=FALSE}
#separate date and time
trailcam_csv$start_date <- as.Date(trailcam_csv$start_time)
trailcam_csv$start_time <- format(trailcam_csv$start_time, "%H:%M:%S")

trailcam_csv$end_date <- as.Date(trailcam_csv$end_time)  
trailcam_csv$end_time <- format(trailcam_csv$end_time, "%H:%M:%S") 
```

---

### Filtering & Joining

```{r Filtering, warning=FALSE, message=FALSE}
#filter for white- tailed deer, select relevant columns, and isolate trail camera id number.
deer_data <- trailcam_csv %>%
  filter(common_name == "White-tailed Deer") %>%
  select(common_name, deployment_id, start_date, start_time, end_date, end_time, group_size) %>%
  mutate(cam_id = as.numeric(str_extract(deployment_id, "\\d+")))
#cam_id refers to the camera id number

#read file with trail camera coordinates
cam_coordinates <- read.csv(here("forest_coordinates.csv"))

#Select id number and coordinates, and rename "No." column to match "cam_id". 
cam_coordinates <- cam_coordinates %>%
  select(No., Longitude, Latitude) %>%
#  rename(cam_id = No.)
#NEED TO FIGURE OUT COORDINATES FOR TRAIL CAMS 
```


```{r Joining, warning=FALSE, message=FALSE}
#join deer data with trail camera coordinate data to find where 
deer_cam_data <- left_join(deer_data, cam_coordinates, by = "cam_id")

#read Duke forest boundary shapefile into project
here("Duke University/Documents/EDE_Fall2023/DukeForest-TackRiosMoyer/duke-forest-spatial-data/duke-forest-spatial-data/Boundary")
forest_sf <- st_read("Duke_Forest_Boundary_Mar2022.shp")
mapview(forest_sf)
```

---

### Coverting to Dataframe

```{r Converting, warning=FALSE, message=FALSE}
#convert coordinates to a spatial dataframe
deer_cam_data_sf <- deer_cam_data_sf %>% st_as_sf(coords = c("Longtitude","Latitude"), crs=4326)
```

---


## Statistical Tests & Machine Learning
Section created by Sophie Moyer

### SUBHEADING 1 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 1 HERE

```


---

### SUBHEADING 2 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 2 HERE

```

---


## Spatial Analysis & Maps
Section created by Grace Randall

### SUBHEADING 1 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 1 HERE

```


---

### SUBHEADING 2 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 2 HERE

```

---

---


## Data Visualization & Plots
Section created by Katie Tack

### SUBHEADING 1 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 1 HERE

```


---

### SUBHEADING 2 HERE

```{r INSERT INFO HERE, warning=FALSE, message=FALSE}
#CODE CHUNK 2 HERE

```

---
